Concrete Semantics for Pushdown Analysis:\\The Essence of Summarization

### 2013-05-09 paper.tex
Added an example.

\usepackage{amsmath,stmaryrd,natbib}
\usepackage[usenames,dvipsnames]{color}

@@

@@

@@

@@

The immediately related work is that of PDCFA
\citet{ianjohnson:earl2010pdcfa},
CFA2~\citet{ianjohnson:vardoulakis-lmcs11,
  ianjohnson:Vardoulakis2011Pushdown}, and AAM
~\citet{dvanhorn:VanHorn2010Abstracting}, the first two of which we
recreated in full detail. The version of CFA2 that handles
\rackett{call/cc} does not handle composable control, and is dependent
on a restricted CPS representation. They also had no way to tune the
precision of their first class continuation approximation. Since
\rackett{call/cc} can be simulated with shift/reset, this work
supercedes theirs.

Pushdown models have existed in the first-order static analysis
literature ~\citet[Chapter
7]{local:muchnick:jones:flow-analysis:1981}\citet{ianjohnson:reps:pushdown:1995},
and the first-order model checking literature
\citet{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time. The
higher-order setting imposes additional challenges that make their
methods difficult to adapt. The most important constraint is that we
can't know all call-sites of a function/method before the analysis
begins, which their methods heavily rely on.

\section{Conclusion}

@@

>>>>

\usepackage{amsmath,stmaryrd,natbib,listings}
\usepackage[usenames,dvipsnames]{color}

@@

\abstract{GAH}

@@

\paragraph{Example:} Let's consider a monovariant allocation strategy
(the address for each binding is the variable name itself) to run the
following example:
\begin{lstlisting}[mathescape]
  (let* ([id ($\lambda$ (x) x)]
         [y (id 0)]
         [z (id 1)])
    ($\le$ y z))
\end{lstlisting}
We should expect that a pushdown analysis would predict this evaluates
to true, and there are no loops in the program. 0CFA claims there is a
loop from the second call of \texttt{id} to the first, and thusly
predicts this program evaluates to true or false. PDCFA claims there
is no loop, and depending on the implementation, that the result is
true or false (paper), or just true (implemented). Let's take a look
at the evaluation after binding \texttt{id} ($\mstore_0 = [\texttt{id} \mapsto \set{\vclo{\slam{\texttt{x}}{\texttt{x}}}{\bot}}]$):

\begin{enumerate}
\item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
    \mstore_0[\texttt{x} \mapsto \set{0}]$ and
    (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
         $\mkont_1$ = \texttt{(letk y (let ([z (id 1)]) ($\le$ y z)) $[\texttt{id}\mapsto{\texttt{id}}]$)})
    $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
\item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
\item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
\item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \texttt{(letk z ($\le$ y z) $[\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]$)} (call this $\mkont_2$).}
\item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
      (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
\item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
\item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
\end{enumerate}

We maintained enough context to distinguish the return points of
\texttt{id} to not rebind \texttt{y} to 1. When determining control
flow through the expression, we consult $\mmemo$ to continue past
function calls, so there is no confusion about a back edge from the
second call to \texttt{id}.

@@

\paragraph{Example:} if we use this semantics to run through the
example of the previous section, we find that with the addition of
stack frames, there is no confusion about \texttt{z}'s value either,
so if the operator weren't $\le$ but instead $<$ or $+$, we could
still constant fold that away.

@@

The immediately related work is that of PDCFA
\citep{ianjohnson:earl2010pdcfa},
CFA2~\citep{ianjohnson:vardoulakis-lmcs11,
  ianjohnson:Vardoulakis2011Pushdown}, and AAM
~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we
recreated in full detail. The version of CFA2 that handles
\rackett{call/cc} does not handle composable control, and is dependent
on a restricted CPS representation. They also had no way to tune the
precision of their first class continuation approximation. Since
\rackett{call/cc} can be simulated with shift/reset, this work
supercedes theirs.

Pushdown models have existed in the first-order static analysis
literature ~\citep[Chapter
7]{local:muchnick:jones:flow-analysis:1981}\citep{ianjohnson:reps:pushdown:1995},
and the first-order model checking literature
\citep{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time. The
higher-order setting imposes additional challenges that make their
methods difficult to adapt. The most important constraint is that we
can't know all call-sites of a function/method before the analysis
begins, which their methods heavily rely on.

\section{Conclusion and future work}

@@

Our goal in the future is to show that this technique is even more
widely applicable than shift and reset, but indeed to the most
intricate control operators in production languages such as in
\citet{ianjohnson:Flatt:2007:ADC:1291151.1291178}. The control
structures there are difficult to model even with AAM's techniques,
but once we can tackle AAM, there should be a straightforward route to
extending it with the pushdown techniques of this paper.

There is also the question of clients of these analyses. Most
obviously we would want to know where we can implement first-class
control more efficiently. On top of optimizations, there are security
analyses. Greater control of the stack gives us the ability to drill
deep into context-sensitive security properties and make precise
predictions.
